
---
title: "Spline Interpolation Conversion: ATQE"
author: "Bryan Roman"
date: "10/28/2024"
output: pdf_document

---

Prep:
  - Working directory
  - Libaries used
  - CSV data that was imported

```{r}
setwd("C:/Users/Roman/OneDrive/Desktop/R_Repo/ATQ_R_Repo")

library(zoo)
library(tidyverse)

# There are two csv files because quarterly interval dates varied
# Debt_To_GDP_Ratio was the outlier 

data = read.csv(file = "Qart_to_Monthly.csv")
data_Debt_To_GDP_Ratio = read.csv(file = "Debt_To_GDP_Ratio.csv")

# Testing one of the variables 
head(data_Debt_To_GDP_Ratio)

```

Created two sub-DataSets
  - data_GDP_Growth
  - data_Account_Balance
  
The negative number removes corresponding column. 
```{r}

data_GDP_Growth = data[,-3]
data_Account_Balance = data[,-2]

# Testing results
head(data_GDP_Growth)
head(data_Account_Balance)

```


This conversion allows use to repeat our time interval in the appropriate manner.

  - Each variable specifies month/date/year for our date
  - There are many ways to read a date, therefor we must specify our format
  

```{r}
# Convert date string to Date object
data$Date = as.Date(data$Date, format = "%m/%d/%Y")
data_GDP_Growth$Date = as.Date(data_GDP_Growth$Date, format = "%m/%d/%Y")
data_Account_Balance$Date = as.Date(data_Account_Balance$Date, format = "%m/%d/%Y")
data_Debt_To_GDP_Ratio$Date = as.Date(data_Debt_To_GDP_Ratio$Date, format = "%m/%d/%Y")

```


#-------------------------------------------------------------------------------

Spline Interpolation function:
  - monthly_dates is our monthly time sequence
  - "Value" is defualt the column name for the data

```{r}
spline_interpolation = function(data, start_date = "1995-01-01", end_date = "2024-01-01") {

    monthly_dates = seq.Date(
    from = as.Date(start_date),
    to = as.Date(end_date),
    by = "month"
  )
  
  x = as.numeric(data$Date)
  y = as.numeric(data[, 2])  
  
  spline_fit = spline(x, y, n = length(monthly_dates))
  
  result = data.frame(
    Date = monthly_dates,
    Value = spline_fit$y
  )
  
  return(result)
}

```

#-------------------------------------------------------------------------------

We rename the default "Value" of each variable

```{r}
monthly_gdp = spline_interpolation(data_GDP_Growth)
names(monthly_gdp)[2] = "GDP_Growth_Rate"

monthly_balance = spline_interpolation(data_Account_Balance)
names(monthly_balance)[2] = "Current_Account_Balance"

monthly_ratio = spline_interpolation(data_Debt_To_GDP_Ratio)
names(monthly_balance)[2] = "Data_Debt_To_GDP_Ratio"
```

Checking results of each variable

```{r}

head(monthly_gdp)
head(monthly_balance)
head(monthly_ratio)

```

Combining the variables into a complete data set
```{r}
complete_data_set = cbind(monthly_gdp,
                      monthly_balance[-1],
                      monthly_ratio[-1]
                      )

# Checking results
head(complete_data_set)
```

Creating a new CSV file to export
```{r}
write.csv(complete_data_set, "Monthly_Interpolated_Data.csv", row.names = FALSE)

```
